Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-07-18 AM+draft.tex
19/07/18 working on discussion v1

For each scenario, we generated 100 replicates of data. A single replicate consists of SNP data and quantitative trait data. 
To introduce some of the complexities of dealing with real genotypes into the study, we obtained the SNP genotypes from 
the publicly available 1000 Genome Project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
by selecting, randomly, a set of SNP loci, assigning additive allelic effects to these loci, and then aggregating these effects for 
each individual along with a random error.  
 The number of SNPs selected per replicate follows a Poisson distribution with mean 30. The sizes of the allelic effects 
 across the selected loci are equal.  A heritability of  50\% is assumed for the trait.

 A replicate was analysed as follows. All programs/packages were run at their default settings. 
 The Eagle package and MLMM script were the easiest of the programs/packages to implement. 
 Their results are immediately 
 interpretable. Their findings are the set of SNPs in strongest association with the trait. Each 
SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.  
Implementing BigRR, LMM-Lasso, and glmnet required more work. 
The computer programs/packages are based on regularisation methods. Being able to fit all the SNPs simultaneously within a regression 
framework is their strength. The analytical challenges of calculating p-values due to the adaptive nature of the estimation procedure is their 
weakness.  To solve this problem, at the cost of increased computation, we calculated the significance of a SNP empirically via stability selection (see Methods). R2VIM is an R package that implements random forests.  
We were required to 
specify the minimum size of a terminal node, the number of trees, and number of potential variables 
for building a tree.

R2VIM implements random forests. As part of analysing data with r2VIM, we were required to specify the minimum size of a terminal node, the number of trees, and number of potential variables 
for building a tree.   HOW TO CHOSE THESE - REASONABLE VALUES ???? LOOK AT PAPERS ...

@@

\subsection{Power and False Discovery Rates}
=============================
How did we define a finding ?????
============================

@@

==>

Of the methods, the findings from Eagle and MLMM are immediately usable. 
Eagle and MLMM, return as their findings
 the set of SNPs in strongest association with the trait. Each 
SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.
For the other methods, FaST-LMM and GEMMA report a P value for each SNP as their primary finding. BigRR, LMM-Lasso, and glmnet, after 
stability selection has been performed (see Methods),  report a probability of importance for each SNP.

FaST-LMM and GEMMA report P values for the statistical significance of a SNP,

For the other methods, either a P value (.....), relative importance measure (r2VIM), or probability of importance (...,, ..., ...) are reported as the 
findings. Before the results can be interpreted, the snp are placed in map order, peak regions (based on the significance measure) are identified, 
a threshold is set, and a set of SNP findings obtained from identifying the SNP with the largest significance measure in those peak regions above the 
threshold.

findings were reported as either a P value (FaST-LMM and GEMMA), importance

Eagle and MLMM, return as their findings
 the set of SNPs in strongest association with the trait. Each 
SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.
BigRR, LMM-Lasso, and glment, after stability selection has been performed (see Methods),

Eagle and MLMM return as their findings the set of SNPs in strongest association with the trait. Each 
SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP. 
The other methods required the identification of SNP in 
BigRR, XX, and XX return as their findings

Interpreting the findings of Eagle and MLMM is trivial. They return as their findings the set of SNPs in strongest association with the trait. Each 
SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP. 
Interpreting the findings of the other methods is more difficult

Only those SNP found to be in strongest association

Eagle and MLMM, return as their findings, those SNPs found to be in strongest association with the trait. Each SNP identifies a separate 
genomic region of interest via it's map location.

The results from Eagle and MLMM required no further processing. Each SNP denotes a separate region of interest where the SNP

To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
in creating the trait data. We then averaged across the 100 replicates.  
Similarly, to calculate the false discovery rate of a method, 
for each replicate, 
we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.  
When a replicate was analysed, 
 a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of a SNP in true 
association with the trait.

<======

When a replicate was analysed, 
 a method's findings were counted as true if the SNPs were located within 40 kilobase pairs of a SNP in true 
association with the trait. To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
in creating the trait data. We then averaged across the 100 replicates.  Similarly, to calculate the false discovery rate of a method, 
for each replicate, 
we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.

@@

@@

>>>>

For each scenario, we generated 100 replicates of data. A single replicate consists of SNP data and quantitative trait data. 
To introduce some of the complexities of dealing with real genotypes into the study, we obtained the SNP genotypes from 
the publicly available 1000 Genome Project, phase 3  \cite{10002010map}.  The quantitative trait data are generated 
by selecting, randomly, a set of SNPs, assigning additive allelic effects to these loci, and then aggregating their effects for 
each individual along with a random error.  
 The number of SNPs selected per replicate follows a Poisson distribution with mean 30. The sizes of the allelic effects 
 across the selected loci are equal.  A heritability of  50\% is assumed for the trait.

 A replicate was analysed as follows. All programs/packages were run at their default settings. 
 Eagle and MLMM were the easiest of the programs/packages to implement. 
 Their results are immediately 
 interpretable. Their findings are the set of SNPs in strongest association with the trait. Each 
SNP in this set identifies a separate genomic region of interest, whose position is given by the map location of the SNP.  
The other programs/packages were not as easy to implement  BigRR, LMM-Lasso, and glmnet make use of regularlisation methods. 
All SNPs are fitted simultaneously in a regression framework.  However, due to the adaptive nature of the estimation procedure, it is challenging, 
analytically, to calculate P values for the effects in the model. 
To solve this problem, at the cost of increased computation, we calculated the significance of a SNP empirically 
via stability selection (see Methods).  R2VIM, a random forest approach, required three parameters to be specified. These are the
 minimum size of a terminal node, the number of trees, and number of potential variables 
for building a tree.  It is not clear how "best" to set these parameters so we followed ref. The minimum size of a node was set to 10\% of the sample size. 
The number of trees was set to 1000. The number of variables was set to 20\% of the number of SNP. A relative importance measure is calculated 
for each SNP measuring the strength of association with the trait. 
FaST-LMM and GEMMA implement single-locus association mapping. The P values of the SNP are reported as their results. 
For all but Eagle and MLMM, the set of SNPs in strongest and significant association with the trait were found 
by placing the SNP in map order, setting a threshold, identifying peak regions with 
significance measures above the threshold, and recording the SNP with the largest significance measure in these peak regions.

@@

\subsection{Power and False Discovery Rates}

@@

To calculate the power of a method, for each replicate, we divided the number of true SNP findings by the number of SNP that had been used 
in creating the trait data. We then averaged across the 100 replicates.  
Similarly, to calculate the false discovery rate of a method, 
for each replicate, 
we divided the number of true SNP findings by the number of (true and false) SNP findings. We then averaged across the 100 replicates.  
A finding was counted as true if the SNP is located within 40 kilobase pairs of a SNP in true 
association with the trait.

@@

Eagle is a new method and R package for multiple locus association mapping. It differs from other multiple locus methods. It can handle large datasets due to dimension reduction It treats association mapping as a model selection instead of a variable selection problem so we don?t have to deal with multiple testing issues. Its results are immediately interpretable. As we saw in the simulation study, Eagle has considerably higher power than single-locus methods and is comparable in run time. However, It was in the analysis of the mouse data that the true worth of Eagle is evident. With Eagle, we found more than double the results than with single locus association mapping, the method of choice and these extra results were all validated as true.

@@

======

General introductory para - gwas growing in size.

Epistatic effects

More complex models

Last paragraph
==============
We see what we are presented here as the first step in the development of Eagle. GWAS are only growing in size, both in terms of number of 
samples and number of genotypes. Our goal here was to develop the methods and implement those methods in package that could be run on a desktop. 
However, the computational power of desktop will not suffice for datasets where there are hundreds of thousands of samples and millions of loci. 
Here, computer clusters will be needed to meet the computational challenges and we have written the Eagle package with this in mind. We do not 
keep the marker data in memory. Also, we perform block multiplication of matrices that would otherwise be too large to fit into memory. These are all 
preemptive steps to make the porting of Eagle onto a cluster environment easier.
